---
title: "Frequentist Concepts"
author: "Rozy Bathrick"
date: "2022-11-23"
output: html_document
---


### Q1 (2 pts.): What is the probability of observing a count of exactly 3 successes in a binomial distribution with parameters n = 4 and p = 0.75?

```{r}
dbinom(3, 4, 0.75)
```

###  Q2 (2 pts.): What is the probability of observing a count of 3 successes or fewer in a binomial distribution with parameters n = 4 and p = 0.75?

```{r}
pbinom(3, 4, 0.75)
```


### Q3 (2 pts.): What is the probability of observing more than 3 successes in a binomial distribution with parameters n = 5 and p = 0.75?

```{r}
1-(pbinom(3, 5, 0.75))
```

### Q4 (2 pts.): - What is the probability of observing a value of less than 1.2 from a normally-distributed population with mean = 2 and standard deviation = 2?

```{r}
pnorm(1.2, mean = 2, sd = 2)
```

### Q5 (2 pts.): - What is the probability of observing a value of greater than 1.2 from a normally-distributed population with mean = 2 and standard deviation = 2?

```{r}
1-(pnorm(1.2, mean=2, sd=2))
```

### Q6 (4 pts.): - What is the probability of observing a value between 1.2 and 3.2 from a normally-distributed population with mean = 2 and standard deviation = 2?

```{r}
pnorm(1.2, 3.2, mean=2, sd=2)
```

### Q7 (2 pts.): Describe how the shape of the histogram changes as you continue to press the sample button

A: The histogram begins to reflect the beta distribution

### Q8 (2 pts.): Describe how the shape of the histogram changes as you continue to press the sample button.

A: The histogram is more narrow, and begins to approach a normal distribution

### Q9 (2 pts.): Describe how the shape of the histogram changes as you continue to press the sample button.

A: The histogram is very narrow, still slightly skewed but closer to normal. 

### Q10 (2 pts.): Why is there such a drastic change in the shape of the sampling distribution when you change the sample size from 1 to 2?

A: because the sample size increases the number of expected values and begins to center observations around the mean.

### Q11 (2 pts.): What are the two main factors that determine the width of the sampling distribution of the mean?

A: standard deviation and sample size.

### Q12 (2 pts.): How many 3-character words are possible?

```{r}
25*25*25
```

### Q13 (2 pts.): How many books would the Library contain if you added one additional position to the book size (i.e. one extra letter on the last page)? Express your answer in terms of B.

B+25^16,4000
