---
title: "A of E D final project"
author: "Rozy Bathrick"
date: "2022-11-21"
output: html_document
---
# Delomys data analysis{.tabset .tabset-pills}

## Data Exploration and Q1-4

First we look at the body length and mass of these little mice in summary and run
a shapiro test to see if the variance is normally distributed


```{r}
require(here)
here()
delo<-read.csv(here("data", "delomys.csv"))
summary(delo$body_mass)
summary(delo$body_length)
shapiro.test(delo$body_mass)
shapiro.test(delo$body_length)
```

### Graphical Exploration

Scatterplot of body length and body width
```{r fig.width=15}

plot(body_length ~ body_mass, data=delo)

```
Histogram of body Mass
```{r}
hist(delo$body_mass, main="Histogram of Delomys body mass", xlab="Body Mass (g)")
```

Histogram of body length
```{r}
hist(delo$body_length, main="Histogram of Delomys body length", xlab="Body Length (mm)")
```
A conditional boxplot of body mass, conditioned on species (column binomial)

```{r}
boxplot(delo$body_mass ~ delo$binomial, main="Body mass of two Delomys Species", xlab=" ", ylab="Body Mass (g)")
```
A conditional boxplot of body mass, conditioned on sex (column binomial)
```{r}
boxplot(delo$body_mass ~ delo$sex, main="Body mass of Delomys Species by sex", xlab=" ", ylab="Body Mass (g)")
```



A conditional boxplot of body mass, conditioned on both species and sex
```{r fig.aspect=20}
boxplot(delo$body_mass ~ delo$sex+ delo$binomial, main="Body mass of two Delomys Species by sex", ylab="Body Mass (g)", xlab=" ", las=2, names=(c("D. dorsalis\n female", "D. dorsalis\nmale", "D. sublineatus\nfemale", "D. sublineatus\nmale")))

```

#### Q1 (2 pts.): Qualitatively describe the relationship between body mass and length.

A: There does seem to be a relationship - it looks logarithmic to me, with diminishing returns in body length as the body mass increases.

#### Q2 (2 pts.): Qualitatively describe the shapes of the histograms.

A: The histograms are unimodal, and appear to be mostly normally distributed with mild skew.

#### Q3 (3 pts.): Using both the histograms and normality tests, do you think the (unconditioned) body masses and body length are normally-distributed?

A: Though the histograms look only slightly asymmetric, the shapiro test returns a low p-value, thus we reject the null that they are normally distributed (they are not normally-distributed) 

#### Q4 (2 pts.): Examine the conditional boxplots. Describe any graphical evidence you see for body mass differences based on species and/or sex.

A: There is not any strong evidence that sex or species are responsible for differences in body mass - the boxplots are mostly overlapping.

## Model Building and Diagnostic Q5-6

```{r}
#Model 1: simple linear regression body_length ~ body_mass
fit1=lm(body_length~body_mass, data=delo)

#Model 2: 1-way ANOVA body_mass ~ sex
fit2=lm(body_mass~sex, data=delo)
           
#Model 3: 1-way ANOVA body_mass ~ binomial
fit3=lm(body_mass~binomial, data=delo)

#Model 4: 2-way additive ANOVA body_mass ~ sex + binomial
fit4=lm(body_mass~sex+binomial, data=delo)

#Model 5: 2-way factorial ANOVA body_mass ~ sex * binomial
fit5=lm(body_mass~sex*binomial, data=delo)
```
```{r}
hist(residuals(fit1))
shapiro.test(residuals(fit1)) #not normally distrib
hist(residuals(fit2))
shapiro.test(residuals(fit2)) #not normally distrib
hist(residuals(fit3))
shapiro.test(residuals(fit3)) #not normally distrib
hist(residuals(fit4))
shapiro.test(residuals(fit4)) #not normally distrib
hist(residuals(fit5))
shapiro.test(residuals(fit5)) #not normally distrib
```

#### Q5 (2 pts.): What do you conclude about residual normality based on the numerical and graphical diagnostics?

A: The residuals are not normally distributed

#### Q6 (1 pt.): Are violations of the normality assumption equally severe for all the models?

A: No. Model 2: 1-way ANOVA explaining body mass by sex is the least severe in its violations, and the first Model is the most severe.


## Model Interpretation Q7-16

```{r}
summary(fit1)
```

#### Q7 (2 pts.): What is the magnitude of the mass/length relationship?

A: For every one mm increase in body length, body mass will increase by 0.875 g

#### Q8 (2 pts.): What is the expected body length of an an animal that weighs 100g?

```{r}
76.12466+(100*.87550)
```


#### Q9 (2 pts.): What is the expected body length of an animal that weighs 0g?
A: 76.12466 mm

```{r}
anova(fit2)
anova(fit3)
anova(fit4)
anova(fit5)
```

#### Q10 (1 pt.): What is the base level for sex?

A: female

#### Q11 (1 pt.): What is the base level for binomial?

A: D. dorsalis

#### Q12 (1 pt.): Which sex is heavier?

A: male

#### Q13 (1 pt.): Which species is heavier?

A: D. dorsalis

#### Q14 (1 pt.): Are sex and species significant predictors for body mass?

A: Yes

#### Q15 (1 pt.): Is there a significant interaction?

A: No

#### Q16 (2 pts.): How dos the significance (as measured by p-value) of either of the main effects (sex and species) differ between the single-predictor models, the additive model, and the interactive model?

A: The significance of the species effect remains constant in all models, at < 2.2e-16 *** (one of the lowest numbers possible). The significance of the sex effect varies a bit, but the p-value remains very small - though never close to as small as the species effect.

## Model Comparison: Body Mass Q17-18


```{r}
##testing the goodness of fit with AIC value

AIC(fit2) #Model 2: 1-way ANOVA body_mass ~ sex
AIC(fit3) #Model 3: 1-way ANOVA body_mass ~ binomial
AIC(fit4) #Model 4: 2-way additive ANOVA body_mass ~ sex + binomial
AIC(fit5) #Model 5: 2-way factorial ANOVA body_mass ~ sex * binomial

```

#### Q17 (1 pt.): Which two models have the lowest AIC?
A: Model 4: 2-way additive ANOVA body_mass ~ sex + binomial and
Model 5: 2-way factorial ANOVA body_mass ~ sex * binomial

#### Q18 (4 pts.): Which of the two models with lowest AIC scores would you select?

Because there isn't a huge difference in AIC between the last two, and we know that the difference between species is really the responsible effect, I would choose Model 4. This is a less complex model, and allows us to see the difference between sex and species effect on body mass.
