---
title: "Probability 3"
author: "Rozy Bathrick"
date: "2022-11-27"
output: html_document
---
### Q1 (1 pt.): If I wanted to use a binomial distribution to model my six forest plots, what values should I use for the two parameters of a binomial distribution?

A: n=6, p=0.667

### Q2 (1 pt.): Use dbinom to calculate the probability of observing birds in exactly four of the six patches. Include your R-code in your answer.

### Q3 (1 pt.): Now, suppose I did a survey and observed no birds in my plots. Use dbinom to calculate the probability of observing no presences.

```{r}
x_bin = 1:6
y_bin_2 = dbinom(x_bin, size = 5, prob = 0.66)

#A2
dbinom(4, 5, 0.66) 

#A3
dbinom(0, 5, 0.66)
```

### Q4 (1 pt.): Back to the binomial scenario (bird presence/absence in 6 forest plots).Now use pbinom to calculate the probability of observing four or fewer presences in the 6 plots. Show your R code.
### Q5 (1 pt.): Now use pbinom and the law of total probability to calculate the probability of observing four or more presences in the 6 plots. Show your R code.

```{r}
#A4
pbinom(4, 5, 0.66)

#A5
1-(pbinom(3, 5, 0.66))

```

### Q6 (1 pt.): Are you more likely to observe a value of 1.0 or 2.0?
A: More likely to observe a 1.0

### Q7 (1 pt.): What is the probability of observing a value of 1.0 or less? Show the R code you used to find your answer.
### Q8 (1 pt.): What is the probability of observing a value between 1.0 and 2.0? Show the R code you used to find your answer.

```{r}
#A7
pnorm(1.0, mean = 0, sd = 1)

#A8
pnorm(2.0, mean=0, sd=1) - pnorm(1.0, mean=0, sd=1)
```
### Q9 (2 pts.): Show the complete R-code you used to create your plot. Make sure you include all the code to recreate your plot in a fresh R session.
### Q10 (1 pt.): Include a figure of your plot.

```{r}
# How many points?
n = 1000

# Create a vector of x-values from -6 to 6:
x = seq(from = -6, to = 6, length.out = n)

# Create the corresponding y-values:
y = dnorm(x, mean = 0, sd = 1)

# plot!
plot(y ~ x, type = "l", ylab = "Probability Density")

# next plot 
y_2 = dnorm(x, mean = 0, sd = 2)

# third plot

y_3 = dnorm(x, mean=-2, sd=1)

plot(y ~ x, type = "l", ylab = "Probability Density")+
points(y_2 ~ x, type = "l", lty  = 2)+
  points(y_3~x, type="b", lty=2)

y_cdf_1 = pnorm(x, mean = 0, sd = 1)
y_cdf_2 = pnorm(x, mean = 0, sd = 2)
y_cdf_3 = pnorm(x, mean=-2, sd=1)
plot(y_cdf_1 ~ x, type = "l", ylab = "Cumulative Density")+
points(y_cdf_2 ~ x, type = "l", lty = 2)+
    points(y_cdf_3~x, type="b", lty=2)

par(mfrow = c(1, 2))
plot(y ~ x, type = "l", ylab = "Probability Density", main="Rozy's PDF Plot")+
points(y_2 ~ x, type = "l", lty  = 2)+
points(y_3~x, type="b", lty=2)
plot(y_cdf_1 ~ x, type = "l", ylab = "Cumulative Density", main="Rozy's CDF Plot")+
points(y_cdf_2 ~ x, type = "l", lty = 2)+
    points(y_cdf_3~x, type="b", lty=2)


```
```


### Q11 (2 pts.): Show the complete R-code you used to create your plot. Make sure you include all the code to recreate your plot in a fresh R session.

### Q12 (1 pt.): Include a figure of your plot.

```{r}

x_bin = 0:6
y_bin_2 = dbinom(x_bin, size = 6, prob = 0.66)

barplot(
  height = y_bin_2,
  # the names to print with each bar:
  names.arg = x_bin,
  # Tells R to remove space between bars:
  space = 0,
  ylab = "Pr(x)",
  main = "Rozy's Binomial: n = 6, p = 2/3")
```


