---
title: "The ultimate guide to battling alongside R"
author: "Rozy Bathrick"
date: "2022-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Reference Guide {.tabset .tabset-pills}


## Loading Data and Packages

First couple of things we'll cover: how do you get some data into R, 
and download the right packages to do what you want with it. 

Loading a pre-existing package: use here and library
```{r}
library(here) #this tells R we are working in the current file pathway
library(palmerpenguins) #this loads the practice dataset of Palmer Penguins (it's already been installed)
require(here) #the require function tells R we want to use it - similar to library
require(palmerpenguins)
```

Next we'll make a data frame out of the ginkgo data we collected as a class

```{r}
require(here)
here()
ginkgo<-read.csv(here("data", "ginkgo_data_2022.csv"))

##here is on the wrong level - its inside repo
```

## Data Structures

The function c() combines its arguments into a vector (a 1-dimensional data structure consisting of 1 or more elements).

All of the elements must be of the same type.
I can’t combine character and numeric types in the same call to c()
Here’s two examples using numeric and character data types:

```{r}
## Create a vector of numbers:
num_vec  = c(1, 4, 8, 9, 13)
## Create a vector of characters:
char_vec = c("there", "is", "a snake" , "in my boot")
char_vec2=c("to", "infinitey", "and", "beyond")
```

I can show the contents of a vector by typing the name of the vector, or using the print() function.

```{r}
## Typing the name of the vector into the console prints the contents
num_vec
## The print() function accomplishes the same task:
print(char_vec)
```

The function length() identifies how many elements are in a vector
```{r}
length(num_vec)
```

The function matrix() creates a grid of vectors that are the same length in elements.
It can also split a single vector into a grid with the dimensions you specify with nrow or byrow
You can also give the rows and columns names with dimnames
```{r}
m1<-matrix(data=char_vec,  nrow=2)
print(m1)
```

The function dataframe()


## Subsetting


## Numerical Data Exploration


## Graphical Data Exploration


## Distribution Functions